\section{Acknowledgments}
We thank Jason Roberts for expert clinical review. We thank Christopher Agnew, Jamie Fraser, James Holton, Lijun Liu, Marco Lolicato, Bruk Mensa, and Tarjani Thaker for technical assistance with crystallographic data collection and processing. We thank Lijun Liu and Tarjani Thaker for technical assistance with protein preparation and crystallization screens. We thank Bruk Mensa for technical assistance with plate reader instrumentation. We thank Michael Grabe, Aimee Kao, Dan Minor, and Oren Rosenberg for helpful discussions. 

Beamline 8.3.1 at the Advanced Light Source is operated by the University of California Office of the President, Multicampus Research Programs and Initiatives grant MR-15-328599, the National Institutes of Health (R01 GM124149 and P30 GM124169), Plexxikon Inc., and the Integrated Diffraction Analysis Technologies program of the US Department of Energy Office of Biological and Environmental Research. The Advanced Light Source (Berkeley, CA) is a national user facility operated by Lawrence Berkeley National Laboratory on behalf of the US Department of Energy under contract number DE-AC02-05CH11231, Office of Basic Energy Sciences. 

E.W.T. was previously supported by a Sarnoff Foundation Fellowship and is currently supported by NIH/NHLBI F30 grant F30HL137329 and NIH/NIGMS grant T32GM007618 to the UCSF Medical Scientist Training Program (MSTP). This work was supported in part by NIH/NHLBI grant DP2HL123228 (to R.C.D.) as well as American Heart Association grant 17IRG33460152 (to R.C.D.).